---
title: "Precardc_Ret_Prog_DTI"
author: "Tasos"
date: "07/10/2022"
output: html_document
---

# Precard analysis for Retinopathy Progression with diabetes Type I

```{r setup,echo=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = FALSE)

library("finalfit")
library("dplyr")
library("survival")
library("knitr")
library('survivalAnalysis')
library("riskRegression")
# library("regplot")
library("rms")
library("pec")
library("survminer")
# library("cmprsk")
# #library("hdnom")
# if (!require("pacman")) install.packages("pacman")
# pacman::p_load(caret, magrittr, pacman, rio, tidyverse)
# #load(file = "Precardc_RET_2021.RData")
```

```{r echo=FALSE, message=FALSE}
html(contents(Ret_prog_DTI), maxlevels=10, levelType='table')
```


# 1. RMS package - Model

## Stepwise backward variable selection for cox ph model
```{r echo=FALSE, message=FALSE, warning=FALSE, comment="",  results='markup'}
# Development of the model using the fastbw() function of the RMS model to identify the optimal combination of variables 
ret.cox_full <- cph(Surv(T2Eg, EventgCox) ~ Age+ Diabetes_Dur+ eGFR+ SYSTOLIC+ DIASTOLIC+ Hba1c+ ACR+
                      Cholesterol+ BMI+ Weight+ Ethnicity_groups+ Afrocaribbean_Ethnicity+ Gender+ ACR_G, 
                    data=Ret_prog_DTI, x=TRUE, y=TRUE, surv=TRUE, time.inc=1)

fastbw(ret.cox_full, rule=c("aic"), type=c("individual"),  sls=.05, aics=1, eps=1e-9, k.aic=2, force=NULL)

```

## Cox regression model

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="",  results='markup'}
# CoxPH models based on the above calculations
ret.cox <- cph(Surv(T2Eg, EventgCox) ~ Afrocaribbean_Ethnicity+SYSTOLIC+Hba1c, 
               data = Ret_prog_DTI, x=TRUE, y=TRUE, surv=TRUE, time.inc=1)
# Display model characteristics
ret.cox

```

## Hazard ratios estimates and 95% CI

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Produce hazard ratios and 95% CIs for model parameters
cat("hazard ratio estimates")
pander(exp(coef(ret.cox))) # hazard ratio estimates)
cat("hazard ratio 95% CI")
pander(exp(confint(ret.cox))) # hazard ratio 95% CI
```

## Checking the Proportional Hazards Assumption

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Proportionality of hazards check via plotting the beta(t) coefficients (line is horizontal) 
cox.zph(ret.cox)
survminer::ggcoxzph(cox.zph(ret.cox), var = c("Afrocaribbean_Ethnicity"), font.main = 12, ylab="Beta (t): Ethnicity", xlab="Follow up (years)")
survminer::ggcoxzph(cox.zph(ret.cox), var = c("SYSTOLIC"), font.main = 12, ylab="Beta (t): Systolic blood pressure", xlab="Follow up (years)")
survminer::ggcoxzph(cox.zph(ret.cox), var = c("Hba1c"), font.main = 12, ylab="Beta (t): Hba1c", xlab="Follow up (years)")
```

## Model evaluation

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}
# Use the anova.rms function for chiq test
anova(ret.cox, test = c( "Chisq"))
```

## Multicollinearity by Variance Inflation Factors

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

cat("Variance Inflation Factors")
pander(rms::vif(ret.cox))

```

## Validation of the model

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Check difference in statistics such as R2, Somers’ d estimate etc, over cross-validated iterations

validate(ret.cox, method="crossvalidation", B=50, bw=FALSE, rule="aic", type="residual", sls=.05, 
         aics=0, force=NULL, estimates=TRUE, pr=FALSE, dxy=TRUE, tol=1e-9)

```

## Calibration plots for 1, 5 and 10 years

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}
# Create calibration plot to check if predicted survival probabilities from the model align with the observed survival probabilities
set.seed(43233)
cal.rms1<-rms::calibrate(ret.cox, cmethod=c('hare', 'KM'), method="boot", u=1, m=150, B=40, 
  bw=TRUE, rule="aic", type="residual", sls=0.05, aics=0, force=NULL,
  estimates=TRUE, pr=FALSE, what="observed-predicted", tol=1e-12, maxdim=5)
cal.rms5<-rms::calibrate(ret.cox5, cmethod=c('hare', 'KM'), method="boot", u=5, m=150, B=40, 
  bw=FALSE, rule="aic", type="residual", sls=0.05, aics=0, force=NULL,
  estimates=TRUE, pr=FALSE, what="observed-predicted", tol=1e-12, maxdim=5)
cal.rms10<-rms::calibrate(ret.cox10, cmethod=c('hare', 'KM'), method="boot", u=10, m=150, B=40, 
  bw=FALSE, rule="aic", type="residual", sls=0.05, aics=0, force=NULL,
  estimates=TRUE, pr=FALSE, what="observed-predicted", tol=1e-12, maxdim=5)
plot(cal.rms1, main="Predicted 1 year Survival", xlab="")
plot(cal.rms5, main="Predicted 5 years Survival", xlab="")
plot(cal.rms10, main="Predicted 10 years Survival", xlab="")

```

## Nomogram of the produced survival model

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup', fig.width=12, fig.asp = .82}

# “Packaging” data before the building of regression model and Nomogram by rms package. This is very important step for Nomogram building. 

dd=datadist(Ret_prog_DTI)
options(datadist="dd")
units(Ret_prog_DTI$T2Eg) <- "Year"

# Nomogram of survival rate will be built based on Cox regression. Set the survival object (surv1, surv2) based on survival function firstly.
surv1<- function(x)surv(1,lp = x) # defined time.inc 1 year survival probability 
surv2<- function(x)surv(5,lp = x) # defined time.inc 5 year survival probability

# Build survival Nomogram of Cox regression 
nom <- nomogram(ret.cox, fun = list(surv1, surv2), funlabel = c("1-year Survival Probability", "5-year Survival Probability"),
                abbrev = TRUE)
plot(nom, xfrac=0.3, cex.axis=.9, nint=20)
```

## PREDICTION with new instances

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Use this prediction function to predict risk at different time points for "new" patient data including the same variables
predictSurvProb(ret.cox, newdata=ret.test.data ,times=c(1, 5, 10, 14))
```

# 2. RiskRegression package

## Proportional hazards and fine-gray models

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Cause specific Cox model
csc <- CSC(Hist(T2Eg, Eventg)~Afrocaribbean_Ethnicity+SYSTOLIC+Hba1c, data = Ret_prog_DTI, cause = "1")
# Print the model characteristics
csc

# Fine gray model
fgr <- FGR(Hist(T2Eg, Eventg)~Afrocaribbean_Ethnicity+SYSTOLIC+Hba1c, data = Ret_prog_DTI, cause = "1")
cat('Competeting Risks model with death as competing Eventg:')
# print the model characteristics
fgr
```

## AUC and Brier plots of the models

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}
# (time-dependent) Brier score and the (time-dependent) AUC in the learning data using bootstrap
# cross-validation. The function optionally provides results for plotting (time-point specific)
##' ROC curves, for (time-point specific) calibration curves and for (time-point specific) retrospective boxplots.
score<-Score(list("Cause-specific Cox"=csc, "Fine-Gray"=fgr), formula = Hist(T2Eg,Eventg)~1,
              data=Ret_prog_DTI, times = seq(1,14,1), plots = "ROC", summary = "risks")
dev.new(width=5,height=4)
plotROC(score,times = 2)
```

## Prediction

### Prediction plot of the absolute risk of variables in a time frame

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}
pred.csc<-predict(csc, newdata = ret.test.data, time = 1:14, cause = 1)
autoplot(pred.csc)
```

### Absolute risk of ONE new patient at a specific time point

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}
# Absolute risk value
predictRisk(csc, newdata = ret.test.data, time = 10, cause = 1)
```

### Absolute risk of multiple new patients at specific time point

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

pfit1 <- predict(csc, newdata = ret.test.data, cause = 1, times = c(14), se = TRUE, keep.newdata = FALSE)
print(pfit1)
```

### Absolute risk alternative plot for 1 patient over 15 and 10 years

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

vec.times <- csc$EventgTimes
pfit1band <- predict(csc, newdata = ret.test.data, cause = 1, times = vec.times, se = TRUE, band = TRUE)

figure3 <- autoplot(pfit1band, band = TRUE, ci = TRUE)$plot
figure3 <- figure3 + xlab("Time (years)") + ylab("Absolute risk") + xlim(0,10)
print(figure3)

```





































































































































































