---
title: "Retinopathy Tables for publication"
author: "Tasos"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("arsenal")
library("DataExplorer")
library("finalfit")
library("dplyr")
library("survival")
library("survminer")
library("knitr")
library("rms")
library(gtsummary)
library(pander)

```


## Flowchart for publication

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.height=6, fig.width=8}

# Flowchart including the number of patients in the various steps of the analysis
DiagrammeR::grViz("digraph {
  graph [layout = dot, rankdir = TB]

  node [shape = rectangle]
  rec1 [label = 'Initial number of patients: 3875']
  rec2 [label = 'Patients with R0M0 baseline values: 2219']
  rec3 [label = 'Final cohort: 1876']
  node [shape = oval]
  rec4 [label = 'Exclude patients with R1M1 or R1M0 baseline values']
  rec5 [label = 'Exclude patients with no follow up']
  # edge definitions with the node IDs
  rec1 -> rec4 -> rec2 -> rec5 -> rec3
  }")

```

\newpage

## Table 1: Patient characteristics at baseline

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Baseline patient characteristics using the gtsummary() function
Ret_prog_DTI%>%
  select(Age, Diabetes_Dur, eGFR, SYSTOLIC, DIASTOLIC, Hba1c, ACR,
         Cholesterol, BMI, Weight, Ethnicity_groups, Afrocaribbean_Ethnicity, Gender, ACR_G)%>%
  tbl_summary(missing = "no", # don't show missing data
              # Define appropriate statistics to display for numerical and categorical data
              statistic = list(all_continuous() ~ "{mean} ({sd}), ({min}, {max}), 
                               {median} ({p25}, {p75})",
                     all_categorical() ~ "{n} ({p}%)"),
              # Define digits to display
              digits = list(all_continuous() ~ c(2, 2)),
              # Change names of variables for better presentation
              label = list(Age ~	"Age (years)",
                           Diabetes_Dur~	"Duration of diabetes",
                           eGFR~	"eGFR baseline",
                           SYSTOLIC ~	"Systolic blood pressure",
                           DIASTOLIC~	"Diastolic blood pressure",
                           Hba1c	~	"HbA1c",
                           ACR ~	"ACR",
                           ACR_G ~	"ACR (groups)",
                           Cholesterol~	"Cholesterol",
                           BMI ~ "BMI", 
                           Afrocaribbean_Ethnicity~	"African Caribbean"	))%>%
  add_n()

```

\newpage

## Table 2: Table with baseline patient characteristics stratified by African-Caribbean Ethnicity

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Same as above, use gtsummary for display of cohort characteristics by ethnicity 
Ret_prog_DTI%>%
  select(Age, Diabetes_Dur, eGFR, SYSTOLIC, DIASTOLIC, Hba1c, ACR,
         Cholesterol, BMI, Weight, Ethnicity_groups, Afrocaribbean_Ethnicity, Gender, ACR_G)%>%
  tbl_summary(missing = "no", by = Afrocaribbean_Ethnicity, # devide data by ethnicity
              statistic = list(all_continuous() ~ "{mean} ({sd}), ({min}, {max}), 
                               {median} ({p25}, {p75})",
                     all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(2, 2)),
              label = list(Age ~	"Age (years)",
                           Diabetes_Dur~	"Duration of diabetes",
                           eGFR~	"eGFR baseline",
                           SYSTOLIC ~	"Systolic blood pressure",
                           DIASTOLIC~	"Diastolic blood pressure",
                           Hba1c	~	"HbA1c",
                           ACR ~	"ACR",
                           ACR_G ~	"ACR (groups)",
                           Cholesterol~	"Cholesterol",
                           BMI ~ "BMI", 
                           Afrocaribbean_Ethnicity~	"African Caribbean"	))%>%
  add_n()%>%
  add_p() # add statistical tests between the groups

```

\newpage

## Retinopathy progression

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Same as above, use gtsummary for display of cohort characteristics by progressors vs non progressors 
Ret_prog_DTI%>%
  select(Age, Diabetes_Dur, eGFR, SYSTOLIC, DIASTOLIC, Hba1c, ACR,
         Cholesterol, BMI, Weight, Ethnicity_groups, EventCox, Afrocaribbean_Ethnicity, Gender, ACR_G)%>%
  # Modify factor levels for better presentation
  mutate(EventCox=factor(EventCox, levels = c(1, 2), 
                         labels=c("progression", "cencored")))%>%
  tbl_summary(missing = "no", by = EventCox,
              statistic = list(all_continuous() ~ "{mean} ({sd}), ({min}, {max}), 
                               {median} ({p25}, {p75})",
                     all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(2, 2)),
              label = list(Age ~	"Age (years)",
                           Diabetes_Dur~	"Duration of diabetes",
                           eGFR~	"eGFR baseline",
                           SYSTOLIC ~	"Systolic blood pressure",
                           DIASTOLIC~	"Diastolic blood pressure",
                           Hba1c	~	"HbA1c",
                           ACR ~	"ACR",
                           ACR_G ~	"ACR (groups)",
                           Cholesterol~	"Cholesterol",
                           BMI ~ "BMI", 
                           Afrocaribbean_Ethnicity~	"African Caribbean"	))%>%
  add_n()%>%
  add_p()

  
```

\newpage

## Retinopathy progression including all cause mortality

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Same as above, use gtsummary for display of cohort characteristics including patients 
# with pre-event death events

Ret_prog_DTI%>%
  select(Age, Diabetes_Dur, eGFR, SYSTOLIC, DIASTOLIC, Hba1c, ACR,
         Cholesterol, BMI, Weight, Ethnicity_groups, Eventg, Afrocaribbean_Ethnicity, Gender, ACR_G)%>%
  mutate(EventCox=factor(Eventg, levels = c(0, 1, 2), 
                         labels=c("cencored", "progression", "pre-event Death")))%>%
  tbl_summary(missing = "no", by = EventCox,
              statistic = list(all_continuous() ~ "{mean} ({sd}), ({min}, {max}), 
                               {median} ({p25}, {p75})",
                     all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(2, 2)),
              label = list(Age ~	"Age (years)",
                           Diabetes_Dur~	"Duration of diabetes",
                           eGFR~	"eGFR baseline",
                           SYSTOLIC ~	"Systolic blood pressure",
                           DIASTOLIC~	"Diastolic blood pressure",
                           Hba1c	~	"HbA1c",
                           ACR ~	"ACR",
                           ACR_G ~	"ACR (groups)",
                           Cholesterol~	"Cholesterol",
                           BMI ~ "BMI", 
                           Afrocaribbean_Ethnicity~	"African Caribbean"	))%>%
  add_n()%>%
  add_p()

```

\newpage

## Regression models

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='asis'}

# Use final fit model for double checking hazard ratios and 95% CIs, 
# also because it is more powerful in presenting all under one table suitable for publication

# Define variable set and survival objects for the models below
explanatory = c("Afrocaribbean_Ethnicity", "SYSTOLIC", "Hba1c")
dependent_dss = "Surv(T2Eg, EventCox)"
dependent_crr = "Surv(T2Eg, Eventg)"

# Create the tables and fill with model outputs
Ret_prog_DTI %>%
  #summary table
  summary_factorlist(dependent_dss, explanatory, column = TRUE, fit_id = TRUE) %>%
    # CPH univariable
  ff_merge(Ret_prog_DTI %>% coxphuni(dependent_dss, explanatory) %>%
     fit2df(estimate_suffix = " (CPH univariable)")) %>%
    # CPH multivariable
  ff_merge(Ret_prog_DTI %>% coxphmulti(dependent_dss, explanatory) %>%
      fit2df(estimate_suffix = " (CPH multivariable)")) %>%
    # Fine and Gray competing risks regression
  ff_merge(Ret_prog_DTI %>% crrmulti(dependent_crr, explanatory) %>%
    fit2df(estimate_suffix = " (Competing risks model)")) %>%
  select(-fit_id, -index) %>%
  dependent_label(Ret_prog_DTI, "Survival")%>%
  kableExtra::kable()

```

\newpage

## Cumulative incidence curve for Afro-caribbean ethnicity

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

fit.test<-survfit(Surv(T2Eg, EventCox) ~ Afrocaribbean_Ethnicity, data = Ret_prog_DTI)

# insert ggplot code

ggsurv_fig1<-ggsurvplot(fit.test,         
          conf.int = F, linetype = "solid", ggtheme = theme_classic(),
          palette = c("#999999", "#E69F00"),
          # Legend specs
          legend.labs=c("Non African-Caribbean", "African-Caribbean"),
          legend.title = "Ethnicity", 
          legend.fontsize=12,
          fun = "cumhaz",
          # axis specs
          xlim=c(0, 12), ylim=c(0, 1),
          font.x = c(14),
          pval = F, font.y = c(14),
          font.tickslab = c(14, "plain"),
          xlab="Follow up (years)",
          ylab="Cumulative Hazard",
          # Table features
          risk.table = T,
          tables.theme = theme_cleantable(),
          tables.y.text = FALSE,
          break.time.by = 2,
          censor.shape="")

# Add additional p value indication in specific position
ggsurv_fig1$plot <- ggsurv_fig1$plot + annotate("text", x = 11.5, y = 0.52, label = "p<0.001", cex=4, col="black", vjust=0, hjust = 1.1, fontface=1.5)

# Save figure in high quality for publication
tiff("Figure_KMplot_Ethnicity.tiff", res = 700, width = 5500, height = 4000, units = "px")
ggsurv_fig1
dev.off()

```

